Causal inference and the detection of coevolutionary contingencies using dynamic phylogenetic models

### 2024-09-05 manuscript.Rmd
Update manuscript.Rmd
Added draft of conclusions and limitations

# Conclusion

xxx

>>>>

# Conclusions and Limitations

First paragraph repeats major advances of the method:
Dynamic model, analogous to Pagel’s discrete, but without being limited to binary traits
Can handle any distribution, latent factors
Can detect evolutionary contingencies (A→B, B→A) and reciprocal causation
Unbiased estimator, empirical examples summary

Second paragraph states limitations, starting with specific to this method then going broader
1: Correlated drift cannot be separated from cross-selection (though this is possible with ‘real’ longitudinal data) - biological examples of traits under correlated drift?
2: Cross-selection effects assumed to be linear - this can be somewhat relaxed with clade-specific parameters, but still locally linear
3: Computational cost - as with any Bayesian model - can range from minutes (for small phylogenies / datasets) to hours (SCCS example) or even days (for the most complex models, or if looped over a sample of trees)
4: Interpretation - we’re detecting contingency, not true causality; though this is the same for other kinds of causal inference (Pearl, Pagel, Turchin)
5: Data quality and inference - comparative data tend to be crude, and all comparative analyses assume that extant variation reflects all past variation. Virtual archeology / paleontology is not as good as real archeology / paleontology! This can be somewhat amended by including information about ancestral nodes, e.g. complexity SCCS analysis put prior for low complexity on last common ancestor - in principle this could be done for any internal node, with prior reflecting exact state of knowledge
